\documentclass[11pt]{article}
\setlength{\textwidth}{150mm}
\setlength{\textheight}{220mm}
\setlength{\oddsidemargin}{10mm}
\setlength{\evensidemargin}{10mm}
\setlength{\topmargin}{-10mm}
\renewcommand{\baselinestretch}{1.3}

\pagestyle{empty}
\begin{document}

\noindent
To:	All Current and Potential Graduate Student Econometrics Lab Users

\noindent
From:	R. Koenker

\noindent
Re:	Request for Proposals

\noindent
Date:	August 1, 1997

\vspace{10mm}
{\it A Little History}:  \hspace{5mm} The econometrics laboratory in Room 458 Commerce West
was funded by a small grant from the UI Ed Tech Board and a much larger
matching grant from the NSF.  The original proposal stated the objective 
of the lab like this:

\vspace{3mm}
\begin{quote}
The objective of the proposal is to enable principal investigators 
to create an Econometrics Lab which would serve the needs of both faculty and
students in econometrics.
\end{quote}

\vspace{3mm}
In practice we have not restricted either faculty or student use of the lab
to those in econometrics but instead have tried to encourage all potential 
demand.  Unfortunately, this policy has gradually led to physical and 
computational congestion and we are in the process of exploring ways to expand the facilities offered by the lab, as well as schemes to divert demand for lab resources.

\vspace{5mm}
{\it A Modest Proposal}: \hspace{5mm}  In an effort to reevaluate demand for the lab and 
assist in planning this expansion we will require, starting with the Fall, 1997 semester,
that all users, {\em including current users}, make a formal proposal to become a lab user.  A one page 
form for the proposal is attached to this memo and will also be available from the lab web site

\vspace{3mm}
\begin{center}
{\tt http://www.econ.uiuc.edu/}$\sim${\tt sosa/eclab.html}
\end{center}

\vspace{3mm}
If you haven't looked at this site recently, you should!  
It contains a wealth of useful links to the web-land of econometrics, and provides further details about lab facilities and resources.
In this proposal you 
will be asked to explain briefly your research problem and why you need/want 
to use the lab.  You should be as explicit as possible about the software
you hope to use and the amount of disk and cpu-time you expect to need 
over the proposal period.  You are encouraged to consult Daniel Morillo 
regarding these aspects of the proposal.  A format for proposals is provided on the following page.

\newpage
\begin{center}
{\bf UI Econometrics Laboratory Proposal}
\end{center}

\vspace{10mm}
Please complete the following form.   {\em Proposals for fall semester use of the lab are  
should be submitted to Professor Koenker} {\bf {\em by September 8th}}.  
Do not exceed space provided on this form.

Evaluation of proposals will be made by a committee of faculty.  
Precedence will be given to econometrics students at the thesis stage, but 
other students with serious computational needs are encouraged to apply.
Questions should be directed to Walter Sosa ({\tt sosa@ragnar.econ.uiuc.edu}) or Daniel Morillo ({\tt morillo@ragnar.econ.uiuc.edu).

\vspace{8mm}
\begin{center}
* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * *
\end{center}

\noindent
{\bf Date}:

\noindent
{\bf Name}:

UI Network ID:

email address:

phone:

econometrics courses (completed):

econometrics courses (currently enrolled):

\vspace{3mm}
\noindent
{\bf Project title}:

Faculty advisor:

\vspace{3mm}
\noindent
{\bf Research project}:\footnote{Focus only on aspect of project which 
requires lab use, in particular, try to explain why other campus computing 
resources are unsatisfactory for the project.}

\vfill
\hfill (over)
\newpage
\noindent
{\bf Software needs}:\footnote{Be as explicit as possible here, for example 
do not simply say ``Splus ,'' say, for example, ``Splus , because only Splus will estimate an 
unbalanced panel model with individual specific heteroscedasticity...''}

\vfill
\noindent
{\bf Disk needs}:\footnote{Estimated disk space required in megabytes.}

\vfill
\noindent
{\bf CPU needs}:\footnote{Expected cpu usage-- try to extrapolate this based
on prior usage and provide some notion of the time path of your proposed 
usage:  will it all be concentrated in November?}

\vfill
\noindent
{\bf Other special pleading}:\footnote{Anything else you feel might be 
{\bf {\em relevant}} to a decision regarding your lab privileges.}
\vfill
\end{document}
