
{\sc Barrodale, I.  and F.D.K. Roberts} (1974).
Solution of an overdetermined system of
equations in the $\ell_1$ norm, {\it Communications ACM,} {\bf 17}, 319-320.

{\sc Buchinsky, M.}, (1998),
Recent Advances in Quantile Regression Models:
A practical guide for empirical research,
J.Human Resources, {\bf 33}, {88-126},


%{\sc Chaudhuri, P.} (1991) Nonparametric estimates of regression quantiles
%and their local Bahadur representation.
%{\it Annals of Statistics},
%{\bf 19}, 760-777.

{\sc Doksum, K.} (1974)  Empirical probability plots and statistical
inference for nonlinear models in the two sample case,
{\it Annals of Statistics}, {\bf 2}, 267-77.
 
{\sc Gutenbrunner, C. and J. \Jureckova} (1992).
Regression quantile and regression rank score process in the linear model and
derived statistics,  {\it Annals of Statistics} {\bf 20}, 305-330.
 
{\sc \Hajek, J. and Z. \Sidak}  (1967).
{\it Theory of Rank Tests},
Academia: Prague.
 
{\sc Hyndman, R.J., D.M Bashtannyk, and G.K. Grunwald}, (1996)
Estimating and Visualizing Conditional Densities, {\it J. of Comp.
and Graphical Stat.} {\bf 5}, 315-36.

{\sc Koenker, R. and G. Bassett} (1978).  Regression quantiles,
{\it Econometrica}, 46, 33-50.

{\sc Koenker, R.} (2001).  Quantile Regression,
{\it J. of Economics Perspectives}, forthcoming.

{\sc Koenker, R.} (2002).  {\it Quantile Regression}, forthcoming.

%{\sc Koenker R.,  P. Ng and S. Portnoy} (1995).
%Quantile Smoothing Splines,
%{\it Biometrika,} {\bf 81}, 673-80.


{\sc Lehmann, E.} (1974)  {\it Nonparametrics:  Statistical Methods Based on
Ranks}, Holden-Day: San Francisco.

{\sc Manski, C.} (1985) Semiparametric analysis of discrete response:
asymptotic properties of the maximum score estimator, {J. of Econometrics},
{\bf 27}, 313-34.

{\sc Portnoy, S. and R. Koenker } (1997).
The Gaussian Hare and the Laplacian Tortoise:
Computability of Squared-error vs. Absolute-error Estimators, with discussion,
{\it Statistical Science}, 12, 279-300.

{\sc Powell, J.L.} (1986) Censored regression quantiles,
{J. of Econometrics}, {\bf 32}, 143-55.

%{\sc Rousseeuw, P. and M. Hubert}, (1998), Regression Depth, with discussion,
%{\it J. Am. Stat.  Assoc.}, {\bf 94}, 388-433.

%{\sc Welsh, A.H.} (1996).
%Robust estimation of smooth regression and spread functions and their
%derivatives, {\it Statistica Sinica}, {\bf 6},347-366.


