\documentclass{article}
\def\RR{\hbox{\raise.4ex\hbox{$\scriptstyle{|}$}\hskip-1.85pt R}}
\pagestyle{empty}
\begin{document}


%\noindent
%University of Illinois \hfill Department of Economics \\
%Spring 1998 \hfill Roger Koenker
\begin{center}
\large{
{ Prospectus for Economics 478 \\
Topics in Semiparametric Econometrics}}
\end{center}

\vspace{3mm}
There has been some student interest expressed in organizing a Spring, 1998
Research Topics in Econometrics course.  
Although I am currently scheduled to teach the undergraduate (371)
econometrics course next semester, I have agreed to prepare a prospectus
for a graduate offering of 478, with the understanding that there will
be a ``market determination'' by the department of which course will
be offered  -- based on comparative demand.  This determination will be made
as I understand it, early in December.  
This course would be organized as an informal seminar.  
There would be no text 
{\em per se}, but there would be a considerable range of readings from the 
current literature.  
There would be no exams but there would be a required term paper,
and there would be periodic student presentations of course material.
The topics covered in the course could be tailored somewhat to student
demand.  I will list a few topics in which I have a research interest,
at the moment, and for which I think there might be some scope for
constructive interaction in a course of this type.

{\bf The meeting time of this course is to be arranged.  The registration 
number for the course is 08052.}


\begin{enumerate}
\item{\bf Open Problems in Quantile Regression}

I would like to devote some time to what I regard as important open research
problems in quantile regression.  These might include some of the following:

\begin{enumerate}
\item  Panel data and random effects models
\item Instrumental variable methods and l1 canonical correlation
\item Recursive estimation in time-series, ARMA, GARCH, etc.
\item Multivariate quantiles and principal curves
\item Rank based inference and related bootstrapping methods.
\end{enumerate}

\item{\bf Comparison of Density and Nonparametric Regression Methods}

The time seems right for another look at the pros and cons of various
nonparametric methods currently available in the literature.  In
particular, I would like to compare current kernel methods with the
sieve-like methods of Gallant and Tauchen's SNP approach and the
log-spline methods of Stone and others.  There are some interesting
current applications to options pricing which might be an interesting
case study here.

\item {\bf Censoring, Sample Selection, and Treatment Effects}

There is a strongly revived interest in the statistical fundementals
of estimating treatment effect models in observational and experimental
studies exemplified by recent work by Rubin, Heckman, Manski and others.

\item {\bf Ecological Regression and Spatial Statistics}

There is an interesting new book by the political scientist Gary King
on the ecological regression fallacy which draws interesting
parallels to current work on the statistical theory of image
processing and tomography, and is also connected to current advances
in spatial statistics. 

\item {\bf Average Derivative Estimation, Single Indices, and the Transformation
Model}

There is still considerable scope for drawing together methods for
estimating transformation models using a variety of techniques 
closely associated with some of the topics listed above.  Here
there are also deeper theoretical issues relating to adaptive estimation.
Survival analysis is an important application in this contextr,
as is the connection to discrete choice models.

\end{enumerate}


\end{document}
