\documentclass{article}
%\input{preamble}
\begin{document}
\bibliographystyle{econometrica}


\noindent
University of Illinois \hfill Department of Economics \\
Spring 2005 \hfill Roger Koenker
\begin{center}
\large{
{ Economics 577 \\
{\bf Topics in Semiparametric Econometrics}}}
\end{center}

\vspace{3mm}
This course is a research seminar in semiparametric methods in
econometrics.  The main objective of the course from a student viewpoint
should be to write a research paper of potentially publishable quality,
either as an exploratory venture directed toward finding a thesis topic,
or as an opportunity to explore a new area in some depth just for the
fun of it.  There will be no exams only a required term paper,

The course will be organized as an informal reading seminar.  
In the first several weeks I plan to work through Devroye and Lugosi (2000).
This will provide a serious introduction to modern empirical process
asymptotics for nonparametric methods.

There will be periodic student presentations of course material.
The topics covered later in the course can be tailored somewhat to student
demand.  There are a few topics in which I have a current research interest,
and for which I think there might be some scope for constructive interaction 
in a course of this type.  I will try to encourage toward these topics
in an effort to increase effective collaboration among students in the
course.  Not surprisingly, some of these topics have some connection
to quantile regression, although in some cases this connection is rather
speculative.  However, initially the focus as in Devroye and Lugosi (2000)
will be on density estimation.  In keeping with my recent interest in
regularization, I hope to spend part of the middle of the course on
penalized maximum likelihood methods for density estimation.

\begin{center}
References
\end{center}
\vspace{5mm}
Devroye, L. and G. Lugosi, (2000) {\it Combinatorial Methods in Density
Estimation}, Springer.

\end{document}
