\input{preamble}
\begin{document}
\bibliographystyle{econometrica}


\noindent
University of Illinois \hfill Department of Economics \\
Spring 1998 \hfill Roger Koenker
\begin{center}
\large{
{ Economics 478 \\
{\bf Topics in Semiparametric Econometrics}}}
\end{center}

\vspace{3mm}
This course is a research seminar in semiparametric methods in
econometrics.  The objective of the course from a student viewpoint
should be to write a research paper of potentially publishable quality,
either as an exploratory venture directed toward finding a thesis topic,
or as an opportunity to explore a new area in some depth just for the
fun of it.

The course will be organized as an informal seminar.  
There will be no text 
{\em per se}, but there will be a considerable range of readings from the 
current literature.  
There will be no exams only a required term paper,
and there will be periodic student presentations of course material.
The topics covered in the course can be tailored somewhat to student
demand.  There are a few topics in which I have a current research interest,
and for which I think there might be some scope for constructive interaction 
in a course of this type.  I will try to encourage toward these topics
in an effort to increase effective collaboration among students in the
course.  Not surprisingly, many of these topics have some connection
to quantile regression, although in some cases this connection is rather
speculative.

In the first class I will try to describe and motivate the topics I have
identified and provide some basic bibliography.

In the second week I would like to get down to business by discussing
a group of recent papers on estimating treatment effects.  This is a
topic on which there is considerable recent activity by both econometricians
and statisticians and I believe that it is worthwhile for the entire class
to take a close look at a few of these papers.  I have prepared a xeroxed
collection of papers which will be distributed through Up Close, 
one of the campus copy centers.  

In the third week I would like to reconsider some classical readings on
instrumental variable methods focusing on the connection between LIML
canonical correlation and some related eigenvalue problems.  The objective
will be to begin to forge a link between the treatment effect discussion
of the previous week and this older literature which would eventually
result in a more refined approach to IV methods for quantile regression.

In the fourth week I will discuss some recent developments on numerical
methods for quantile regression, and raise some questions about my
ongoing project to revise the software distribution for Splus for
quantile regression.  This discussion is likely to focus on extensions
of existing methods for inference.

In subsequent weeks there will be a sequence of mini-lectures by students
to report progress on their research.  Since the scope for these
presentations will be rather severely constrained by the size of the
class, I am investigating schemes which would enable students to ``present''
research results on lab web pages.  This would enable us to share bibliography,
software and other materials in an efficient and timely manner.  I hope
to provide some tutorial material on this in the first weeks of class.
to provide some tutorial material on this in the first weeks of class.
Some basic material related to the course, including this document, is
available on the web at
{\tt http://www.econ.uiuc.edu/~roger/courses/478/478.html}.
There will also be an expanded version of the bibliography 
which appears below which will
be available on the web.  One of the byproducts of the course
will be an expanded version of this bibliography. 




I would strongly recommend that you to begin to 
learn \LaTeX and Splus as part of
the research process. Some (hopefully helpful) hints on this are
available in my paper ``Reproducible Econometrics'' which I will distribute
in class.  
The book by \cite{VR:1997}  is highly recommended
for Splus.  \cite{Lamp:1994} is the definitive introduction to \LaTeX.



\begin{enumerate}

\item {\bf Endogoneity, Sample Selection, and Treatment Effects}

There is a strongly revived interest in the statistical fundementals
of estimating treatment effect models in observational and experimental
studies exemplified by recent work by 
\cite{air:1996,aai:1997,heckman:1996,ai:1994,Manski:1997}.
This will be our first topic and I think that it will provide a useful
fresh look at the basic idea of instrumental variables
which is one of the few, really original, ideas in econometrics.
After reading some of these new papers, I would like to take a week
to revisit some of the classical literature on IV's, LIML and related
topics from the 1950's.  This literature will, I hope, lead us into
a deeper consideration of how to apply these ideas to the next topic.


\item{\bf Open Problems in Quantile Regression}

I would like to devote some time to what I regard as important open research
problems in quantile regression.  These might include some of the following:

\begin{itemize}
\item Computational methods for quantile regression 
\cite{PK.1997,K.1997,P.1997,F.1997,KP.96}
\item  Panel data and random effects models 
\cite{Chay,Twins,rock:1983,mark:horo:1995}
\item Instrumental variable methods and l1 canonical correlation 
\cite{Amem:82,P:83,amma.93,Twins}
\item Influence robust quantile regression \cite{RH:97,HP:97,JWW:88}
\item Recursive estimation in time-series, ARMA, GARCH, etc. 
\cite{KZ.96,hols:1985,hols:1987,engl:hols:rupp:1988,rupp:1985}
\item Binary Response Models with Heteroscedasticity and
Manski's Maximum Score Estimator \cite{M.75,M.85,mt:86,horo}
\item Multivariate quantiles and principal curves 
\cite{Oja.83,NON,oja:niin:1985,niin:oja:tabl:1990,niin:oja:1995,CC,Chak,Chd.96,Liu,RS,Small:1990,Small,kolt,hast:stue:1989}
\item Rank based inference and related bootstrapping methods 
\cite{K.94,K.96,B.94,Bu.95,Hahn,GJ.92,GJKP.94,horo:97,parz:wei:ying:1994,Fitz.96,G.94,KM.98}
\end{itemize}

\item{\bf Comparison of Density and Nonparametric Regression Methods}

The time seems right for another look at the pros and cons of various
nonparametric methods currently available in the literature.  In
particular, I would like to compare current kernel methods with the
sieve-like methods of the \cite{FG:1996,GT:1997} SNP approach and the
log-spline methods of \cite{SHKT} and others.  There are some interesting
current applications to options pricing which might be an interesting
case study here.


\item {\bf Ecological Regression and Spatial Statistics}

There is an interesting new book by a political scientist \cite{King}
on the ecological regression fallacy which draws interesting
parallels to current work on the statistical theory of image
processing and tomography, and is also connected to current advances
in spatial statistics, see for example \cite{Small}.
There are interesting connections to density estimation, nonparametric
quantile regression and other topics which are suggested by these
developments.

\item {\bf Average Derivative Estimation, Single Indices, 
and the Transformation Model}

There is still considerable scope for drawing together methods for
estimating transformation models using a variety of techniques 
closely associated with some of the topics listed above.  Here
there are also deeper theoretical issues relating to adaptive estimation.
Single index models have a large literature in econometrics, see e.g.
\cite{Stoker,cds:1997} 
Survival analysis is an important application in this context,
as is the connection to discrete choice models.
There are also interesting connections to rank statistics, see 
\cite{Han:1987,G:1997}.

\end{enumerate}


\bibliography{478s98}
\end{document}
